---
title: "esm_244_gorchels"
author: "Madeline Gorchels"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:
- basic review (data wrangling)
- intro to spatial data in R with sf
- intro to shiny apps

1.Loading Packages
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(tmap)
```

2. Review in reading, wrangling and exploring data 

a. Read in the data
```{r}
wine = read_csv("wine_data.csv")

```

b. Wrangling
```{r}
wine_new = wine %>%
  select(country, province, region_1, points, price, winery) %>% 
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd)
  
```

c. Use group_by and summarize to find summary statistics by group

```{r}

wine_summary = wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), 
    max_ppd = max(ppd, na.rm = TRUE)
    
  )

```

d. A few gg-graphics

```{r}
wine_hists = ggplot(wine_new, aes(x = ppd))+
  geom_histogram(aes(fill = state))+
  theme_classic()+
  facet_wrap(~state, scales = "free")

wine_hists
```
new graph: scatter plot of price vs points adding a linear model trendline on top of it 

```{r}
wine_points = ggplot(wine_new, aes(x = price, y = points))+
  geom_point(aes(color = state), alpha = 0.5)+
  geom_smooth(color = "black", method = "lm", se = FALSE)+
  theme_classic()+
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange"))+
  facet_wrap(~state)
  

wine_points

```

2. California Counties Spatial Data Using sf

```{r}
ca_counties = st_read(dsn = ".", layer = "CA_Counties_TIGER2016")

# plot(ca_counties)

ca_land = ca_counties %>% 
  select(NAME, ALAND) 
 #can't currently match because the variables are named NAMES and COUNTY
ca_pop_inc = read_csv("ca_pop_inc.csv") %>%
  rename(NAME = COUNTY)

ca_df = full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome) #geometries will be rememebered


```

Now lets use geom_sf to make a map

```{r}
ca_income = ggplot(ca_df)+
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

#leaflet(ca_df) %>% #need to change the coordinate reference system to match
  #addPolygons()
ca_df_transform = st_transform(ca_df, crs = 4326)
leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()
```

Now lets make an app

```{r}

```

